Escape Analysis and Specialization in a Tracing JIT

### 2010-09-28 talk/pepm2011/paper.tex
fix an XXX

This process is called \emph{escape analysis}. The escape analysis of
our tracing JIT works by using \emph{virtual objects}: The trace is walked from
beginning to end and whenever a \texttt{new} operation is seen, the operation is
removed and a virtual object\arigo{XXX what I have in mind when I talk
of ``virtual object'' is the run-time behavior -- i.e. a real object that
would exist at run-time, except that it has be virtual-ized.  Here you seem
to mean rather ``virtual object description'' or something.}
is constructed. The virtual object summarizes the
shape of the object that is allocated at this position in the original trace,
and is used by the optimization to improve the trace. The shapes describe
where the values that would be stored in the fields of the allocated objects
come from. Whenever the optimizer sees a \texttt{setfield} that writes into a virtual
object, that shape summary is thus updated and the operation can be removed.
When the optimizer encounters a \texttt{getfield} from a virtual, the result is read
from the virtual object, and the operation is also removed.

XXX what happens on a guard\_class?

@@

So far we have only described what happens when virtual objects are used in
operations that read and write their fields. When the virtual object is used in
any other operation, it cannot stay virtual. For example, when a virtual object
is stored in a globally accessible place, the object needs to actually be
allocated, as it will live longer than one iteration of the loop.

>>>>

This process is called \emph{escape analysis}. The escape analysis of
our tracing JIT works by using \emph{virtual objects}: The trace is walked from
beginning to end and whenever a \texttt{new} operation is seen, the operation is
removed and a virtual object\arigo{XXX what I have in mind when I talk
of ``virtual object'' is the run-time behavior -- i.e. a real object that
would exist at run-time, except that it has be virtual-ized.  Here you seem
to mean rather ``virtual object description'' or something.}
is constructed. The virtual object summarizes the
shape of the object that is allocated at this position in the original trace,
and is used by the optimization to improve the trace. The shapes describe
where the values that would be stored in the fields of the allocated objects
come from. Whenever the optimizer sees a \texttt{setfield} that writes into a virtual
object, that shape summary is thus updated and the operation can be removed.
When the optimizer encounters a \texttt{getfield} from a virtual, the result is read
from the virtual object, and the operation is also removed.
Equivalently, a \texttt{guard\_class} on a virtual object can be removed as
well, because the virtual object has a fixed and known class.

@@

So far we have only described what happens when virtual objects are used in
operations that read and write their fields and in guards. When the virtual
object is used in
any other operation, it cannot stay virtual. For example, when a virtual object
is stored in a globally accessible place, the object needs to actually be
allocated, as it might live longer than one iteration of the loop.
